---
title: "Explore BpSs-historic"
author: "Randy Swaty"
date: "2023-03-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## To do

* get sum of acres
* get acres open vs. closed
* bps bar
* stacked bar with percent of old being open or closed
* stacked bar with open old, closed old and early, mid



## Set up: set options, load packages and read in data

```{r setup, message=FALSE, warning=FALSE}
#prevent scientific notation
options(scipen = 999999)

# load packages
library(scales)
library(stringr)
library(tidyverse)

# read in data

bps_mz2 <- read_csv("data/bps_mz2.csv")

bps_mz2_clean <- read_csv("data/historic_bps_clean.csv")

head(bps_mz2)
```

## Try Bar Chart

```{r bps bar chart, fig.height= 15, fig.width=9.5, fig.align='left'}

bps_mz2$shortName <- str_wrap(bps_mz2$bps_name, width = 30)


bpsChart <- 
  ggplot(data = bps_mz2, aes(x = reorder(shortName,acres), y = acres)) +
  geom_bar(stat = "identity") +
  labs(
    title = "",
    subtitle = "",
    caption = "",
    x = "",
    y = "Acres") +
  coord_flip(clip = "off") +
  theme_bw() +
  scale_y_continuous(labels = comma,
                     limits = c(-1000, 2500000)) 

bpsChart

```


## Stacked bar to explore percent of old that is open vs. closed


DATA IS WRONG AS OF 21 MARCH 2023

```{r open closed old}

bps_mz2_long <- bps_mz2 %>%
  select(-c(1,2,5)) %>%
  pivot_longer(!shortName, names_to = "canopy", values_to = "percentOld")


open_closed_chart <-
  ggplot(data = bps_mz2_long, 
         aes(x = shortName, 
             y = percentOld, 
             fill = canopy)) +
  geom_bar(position = "stack", stat = "identity") +
  labs(
    title = "",
    subtitle = "",
    caption = "",
    x = "",
    y = "Acres") +
  coord_flip(clip = "off") +
  theme_bw() +
  # scale_fill_manual(values = c("#a7dbb3", 
  #                              "#3f5243")) +
 scale_x_discrete(limits = rev) + 
  scale_fill_manual(values=c("#3f5243", "#a7dbb3"), 
                       name="Amount of old \nsuccession classes with\nOpen or Closed canopy",
                       breaks=c("percentClosed", "percentOpen"),
                       labels=c("Closed", "Open"))


open_closed_chart

```



## Stacked bar to explore old-open vs. old-closed vs. early-mid

DATA IS WRONG AS OF 21 MARCH 2023

```{r open closed old mid early, fig.height= 15, fig.width=9.5, fig.align='left'}

bps_mz2_clean$shortName <- str_wrap(bps_mz2_clean$BpS_Name, width = 30)

bps_mz2_clean_long <- bps_mz2_clean %>%
  select(-c(1:5)) %>%
  pivot_longer(!shortName, names_to = "label", values_to = "percent")


oo_oc_em_chart <-
  ggplot(data = bps_mz2_clean_long, 
         aes(x = shortName, 
             y = percent, 
             fill = label)) +
  geom_bar(position = "stack", stat = "identity") +
  labs(
    title = "",
    subtitle = "",
    caption = "",
    x = "",
    y = "Acres") +
  coord_flip(clip = "off") +
  theme_bw() +
  # scale_fill_manual(values = c("#a7dbb3", 
  #                              "#3f5243")) +
 scale_x_discrete(limits = rev) + 
  scale_fill_manual(values=c("#3f5243", "#a7dbb3", "red"), 
                       name="Amount of old-closed \nvs. old-open \nvs. early and mid",
                       breaks=c("percent of bps old-closed", "percent of bps old-open", "percent of bps-early or mid"),
                       labels=c("Closed-Old", "Open-Old", "Early and mid"))


oo_oc_em_chart

```

